---
title: "SOEP-IS Workshop"
author: "Lisa Reiber"
date: '2020-11-03'
output:
  html_document:
    code_folding: show
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, warning = TRUE) 

#' show two significant digits tops
options(digits = 2)
#' tend not to show scientific notation, because we're just psychologists
options(scipen = 7)
#' make output a bit wider
options(width = 190)
#' set a seed to make analyses depending on random number generation reproducible
set.seed(2411) # if you use your significant other's birthday make sure you stay together for the sake of reproducibility

source(here::here("00_packages.R"))
source(here::here("00_functions.R"))

# Tidylog off = list(), tidylog on = NULL
# options("tidylog.display" = NULL)
tidylog_off()
```

# Overall Goal

Join information from different SOEP-IS datasets and survey years. 

1. Import SOEP-IS data and generate sample of interest.
2. Find additional variables of interest (Big 5)
3. Join everything together.

Extra:

- How to work with tidylog
- How to work with labelled data
- How to work reproducible

# Import SOEP Data

## set the paths to confidential data 

this step needs to be changed by hand for every local computer. 
Define the relative paths to the data, from wherever you saved the .Rproj file.
```{r}
path_soep <- here::here("../../data/00_raw/SOEP/")

path_soepisv36 <- paste0(path_soep, "SOEP_v36/Gene/")
path_soepisv34 <- paste0(path_soep, "SOEP_v34/soep-is.2017_stata_en/soep-is.2017_stata_en/")
```

# Define sample of interest

- What is our sample of interest?

Our sample of interests are the participants of the SOEP-IS v36 sample with valid saliva probes.


## Import igene sample
```{r}
soepis_igene_raw <- import_soep("Inno19F_IGENE.dta", path_soepisv36)
```

check it out
```{r}
# soepis_igene_raw %>% glimpse() 
# soepis_igene_raw %>% View() 
# soepis_igene_raw %>% head() %>% str() 
soepis_igene_raw %>% head()

```

Where are the value labels?
In the attributes. Here is an example on how to get variable and value labels of the igene06 variable
```{r}
soepis_igene_raw %>% select(igene06) %>% str 
soepis_igene_raw %>% select(igene06) %>% sjlabelled::get_label() 
soepis_igene_raw %>% select(igene06) %>% sjlabelled::get_labels()
```

### Exercise Labels: 
- What is the variable Label of the last variable in the dataset?
- What are the value labels?

## Generate Labelled Dataset
```{r}
igene_lab <- soepis_igene_raw %>% mutate_all(sjlabelled::as_label) 

# generate a list of frequency tables for the show_table function
tables_list <- purrr::map(igene_lab, janitor::tabyl)
tables_list_l <- tables_list
names(tables_list_l) <- purrr::map_chr(igene_lab, get_label) 
```

check out the differnence: Example for variable igene06 
```{r}
# soepis_igene_raw %>% select(igene06) %>% View()
# soepis_igene_raw %>% select(igene06) %>% str()
# soepis_igene_raw %>% tabyl(igene06)
soepis_igene_raw %>% count(igene06)
```

We want to know what the variable stands for and what the value labels are:
```{r}
# igene_lab %>% show_table("igene06", tidy = FALSE)
igene_lab %>% show_table("igene06", tidy = TRUE)
```

### Exercise Data Exploration:
- Get a feel for the dataset and look into a few other Variables:
- How would you describe the format of the data (wide or long?)
```{r}

```

## Filter for valid indivdiuals
```{r}
igene_sample <- igene_lab %>% 
      filter(probe == "ans Labor geschickt") %>% 
      mutate(pid = as_numeric(pid))
```

# Import and subset ppfad data

ppfad (Person-related Meta-dataset) is the most extensive dataset, including most peole who ever participated in the SOEP-IS. It is therefore used as the “base” dataset, onto which the other information will be merged. [ppfad documentation](https://paneldata.org/soep-is/data/ppfad)
```{r}
ppfad_wide <- import_soep("ppfad.dta") %>% 
    select(# time invariant
           pid,   
           cid,
           psample,
           sex,
           gebjahr,
           gebmonat,
           # time variant
           starts_with("hid"),
           contains("netto")
           ) 

ppfad_igene_wide <- ppfad_wide %>% 
  # filter to keep only individuals from igene_sample
  semi_join(igene_sample %>% distinct(pid), by = "pid") 
```
this format is in wide format, but we want it in long  format

wide: one row per individual (pid)
long: muliple rows per individual, unique rows per individual and year (pid, syear)

```{r}
ppfad_igene_wide %>% names
```

```{r}
time_invariant_vars <- c("pid", "cid", "psample", "sex", "gebjahr", "gebmonat")

ppfad_labels <- ppfad_igene_wide %>% select(netto = netto00)

ppfad_igene_long <- ppfad_igene_wide %>% 
         pivot_longer(cols = -all_of(time_invariant_vars), 
                      names_to = "var_syear", 
                      values_to = "value") %>% 
         tidyr::separate(var_syear, into = c("var", "syear"), sep = -2) %>% 
         mutate(syear = ifelse(syear < 40, paste0("20", syear), paste0("19", syear))) %>% 
         pivot_wider(names_from = "var", 
                     values_from = "value") %>% 
         codebook::rescue_attributes(ppfad_labels) %>% 
         mutate(syear = as.numeric(syear),
                netto_l = sjlabelled::as_label(.$netto),
                sex_l = sjlabelled::as_label(.$sex)) %>% 
         sjlabelled::var_labels(syear = "Survey Year",
                             netto = "Survey Status",
                             hid = "Household Number") 

#simulate v35 and v36 (if you work with the newest version, you dont need that)
simulate_waves = TRUE

if (simulate_waves == TRUE) {
    ppfad_v35 <- ppfad_igene_long %>% 
                  select(all_of(time_invariant_vars)) %>% 
                  distinct(pid, .keep_all = TRUE) %>%
                  mutate(syear = 2018)
    
    ppfad_v36 <- ppfad_igene_long %>% 
                  select(all_of(time_invariant_vars)) %>% 
                  distinct(pid, .keep_all = TRUE) %>%
                  mutate(syear = 2019)
                  
    ppfad_igene_long <- ppfad_igene_long %>% 
        bind_rows(ppfad_v35) %>% 
        bind_rows(ppfad_v36) %>%
        rescue_attributes(ppfad_igene_long)
}
```

```{r}
ppfad_igene_long %>% names
# ppfad_igene_long %>% tabyl(syear)
```


# Find big 5

Now we want to add the available information on the big 5 to our sample of interest.

## Exercise:

- Generate a dataset that contains the information on the big 5 `big5_igene`
- join the ppfad data and big5 data into one dataset
  - watch out for the structure of the datasets (identifier keys)
  
Hint:
- the p [Variables from the Individual Question Module](https://paneldata.org/soep-is/data/p) dataset contains information from the person individual questionnaire on the person level, which all adults get (no children)
- [Psychological Scales Manual SOEP](https://www.diw.de/sixcms/detail.php?id=diw_01.c.554370.de)


```{r, child = "workshop_solutions.Rmd"}
```


# Show Distributions

We starts with ppfad and add igene data, then big 5 `big5_igene`

```{r}
tidylog_on()

igene_data <- ppfad_igene_long %>% 
  left_join(big5_igene, by = c("pid", "syear")) %>% 
  codebook::rescue_attributes(ppfad_igene_long) %>% 
  codebook::rescue_attributes(big5_igene)

igene_data_cl <- igene_data %>%
  codebook::detect_missing(learn_from_labels = T, 
                           negative_values_are_missing = T, 
                           only_labelled = F)
```

```{r}
tidylog_off()

igene_data_l <- igene_data_cl %>% mutate_all(sjlabelled::as_label)

# generate a list of frequency tables for the show_table function
tables_list <- purrr::map(igene_data_l, ~janitor::tabyl(.x, show_missing_levels = FALSE))
tables_list_l <- tables_list
names(tables_list_l) <- purrr::map_chr(igene_data_l, sjlabelled::get_label) 
```

```{r}
igene_data_l %>% tabyl(syear, b5_open_2, show_missing_levels = F)

igene_data_l %>% tabyl(b5_open_1) 
```

```{r}
igene_data_l %>% tabyl(syear, b5_open_3, show_missing_levels = F)
```



