# Solutions

## Exercise Labels: 
- What is the variable Label of the last variable in the dataset?
- What are the value labels?

```{r}
soepis_igene_raw %>% select(ncol(.)) %>% get_label
soepis_igene_raw %>% select(length(soepis_igene_raw)) %>% get_labels
```

## Exercise Data Exploration:
- Get a feel for the dataset and look into a few other Variables:
- How would you describe the format of the data (wide or long?)
```{r}
igene_lab %>% select(pid) %>% n_distinct()
igene_lab %>% select(pid, syear) %>% n_distinct()
igene_lab %>% count(syear)
```

## Exercise Big 5:

- Generate a dataset that contains the information on the big 5
- join the ppfad data and big5 data into one dataset
  - watch out for the structure of the datasets (identifier keys)

Hint:
- the p [Variables from the Individual Question Module](https://paneldata.org/soep-is/data/p) dataset contains information from the person individual questionnaire on the person level, which all adults get (no children)
- [Psychological Scales Manual SOEP](https://www.diw.de/sixcms/detail.php?id=diw_01.c.554370.de)

```{r}
p_igene <- import_soep("p.dta") %>% 
  semi_join(ppfad_igene_long %>% distinct(pid), by = "pid")
```

```{r}
big5_igene <- p_igene %>% 
  select(
    # ID's
    pid, syear,
    # Personality
    b5_open_1 = plh0212,
    b5_open_2 = plh0215,
    b5_open_3 = plh0220,
    b5_open_4 = plh0225,
    b5_consc_1 = plh0218,
    b5_consc_2 = plh0222,
    b5_extra_1 = plh0213,
    b5_extra_2 = plh0219,
    b5_extra_3 = plh0223,
    b5_agree_1 = plh0214,
    b5_agree_2 = plh0217,
    b5_agree_3 = plh0224,
    b5_neuro_1 = plh0216,
    b5_neuro_2 = plh0221,
    b5_neuro_3 = plh0226
    ) %>% 
  codebook::rescue_attributes(p_igene) %>% 
  codebook::detect_missing(learn_from_labels = T, 
                                negative_values_are_missing = T, 
                                ninety_nine_problems = F, 
                                only_labelled = F) 
```

## Merge Multiple Datasets

We starts with ppfad and add igene data, then big 5

```{r}
tidylog_on()

igene_data <- ppfad_igene_long %>% 
  left_join(big5_igene, by = c("pid", "syear")) %>% 
  codebook::rescue_attributes(ppfad_igene_long) %>% 
  codebook::rescue_attributes(big5_igene)
```



